\documentclass[upright, contnum]{umemoria}
\depto{DEPARTAMENTO DE INGENIERÍA ELÉCTRICA}
\author{FRANCO ANDREAS CUROTTO MOLINA}
\title{GRAPHSLAM ALGORITHM IMPLEMENTATION FOR sOLVING THE PROBLEM OF SLAM}
\auspicio{}
\date{DICIEMBRE 2015}
\guia{MARTIN ADAMS}
\carrera{INGENIERO CIVIL ELÉCTRICO}
\memoria{MEMORIA PARA OPTAR AL TÍTULO DE}
\comision{\ Marcos Eduardo Orchard Concha}{\ Jorge Felipe Silva Sánchez} %{\ }

\usepackage{lipsum}

\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}

\input{packages.tex}
\newcommand{\estWidth}{0.8}
\newcommand{\errorWidth}{0.5}


\begin{document}

\frontmatter
\maketitle

\begin{abstract}
SLAM (Simultaneous Localization and Mapping) is the problem of concurrently estimate a robot (or other agent) position and the map the robot is generating of the environment. It is consider to be a core concept in mobile robotics, and fundamental to achieve truly autonomous systems. 

Among the many solution that has been proposed for solving SLAM, graph based approaches have gain significant interest from researches in the recent years. These solutions presents various advantages, such as, the capability to handle larges amount of data, and to retrieve the whole robot trajectory, rather than just the last position. A particular implementation of this approach is the GraphSLAM algorithm, first presented Thrun and Montemerlo in 2006.

In the present work a GraphSLAM algorithm is implemented for solving the SLAM problem in a two dimensional scenario. The main objective of this work is to provide of a widely accepted SLAM solution for benchmark comparison for newer SLAM algorithms. The implementation uses the g$^2$o framework as a tool for nonlinear least square optimization.  The algorithm utilize a sparse matrix to store the robot uncertainty in an efficient manner. 

The GraphSLAM implementation is able to solve SLAM with known and unknown data association. This means that even with the robot is incapable to distinguish two measurements belonging to the same landmark, it can assert this its equivalence by computing the likelihood of the landmarks being the same. The algorithm also uses a kernel method for robust estimation against outliers. In order to improve the algorithm computational time, several strategies were designed to efficiently test the association between landmarks, and run the optimizations. 

The final implementation was tested with simulated and real data, in the case of known and unknown data association. It worked successfully in all the tasted cases, being able to retrieved the robot path and the environment map with a low error. The main advantage of the algorithm are the high precision and the available number of parameters to accommodate it to a wide range of scenarios. The mayor drawbacks are the algorithm computational time for large dataset, and the inability to deal with false alarms. 

Finally, future work is given in order to improve the implementation current performance. In particular, modifications are suggested to increase convergence speed, and a method for dealing with false positives is proposed.   
\end{abstract}

\begin{dedicatoria} % opcional
Una dedicatoria corta. Por ejemplo, \emph{A los creadores de U-Campus}
\end{dedicatoria}

\begin{thanks} % opcional
%\lipsum[1-2]
\end{thanks}
\cleardoublepage

\tableofcontents
\listoftables % opcional
\listoffigures % opcional

\mainmatter

%\input{intro.tex}
\input{cap1.tex}
\input{cap2.tex}
\input{cap3.tex}
\input{cap4.tex}
\input{conclu.tex}

%\input{glosario.tex} % opcional

\bibliographystyle{plain}
\bibliography{bibliografia}

% \input{anexo_apendices.tex} % opcionales

\end{document}
